---
title: Exploratory Data Analyst (SQL)
author: 'lynne '
date: '2022-06-21'
slug: exploratory-data-analyst-sql
categories:
  - learn
tags:
  - R Markdown
---


# Exploratory Data Analyst applies to the following cases:

1. vague targets or problems to the large data set 

2. large data set needs cleaning 

# Benefit of the exploratory data analyst:

1. help find the inner pattern or relationship through the data 

2. help to categorize the variables for the further analysis. (missing, outlier etc.)

# Three type of data 

## Numeric data

We would like to know the min and max, average, variance and correlation and percentile to this type of data.  
Need to pay attention to learn how to truncate and round numeric values if necessary.  
create temporary tables to save the interested variable in order to run the correlation matrix.  

```{r, eval = F}

select min(), max(), stddev(), avg()

from table 

# median is different from other statistics as it need to be sorted by self or other linked variable 
percentile_disc(0.5) 
WITHIN GROUP (ORDER BY column_name)

```



```{r, eval = F}


SELECT TRUNCATE(123.4567, 3);   # 123.456
SELECT TRUNCATE(123.4567, 2);   # 123.45
SELECT TRUNCATE(123.4567, 1);   # 123.4
SELECT TRUNCATE(123.4567, 0);   # 123
SELECT TRUNCATE(123.4567, -1);  # 120
SELECT TRUNCATE(123.4567, -2);  # 100
SELECT TRUNCATE(123.4567, -3);  # 0

```


## Catergorical data or unstructure data 

un-structure data is useful if we structure it into an appropriate way.  
Need to deal with case, spacing and delimiters (special characterics)
create a temporary table to count and aggregate any interested variable.  

```{r, eval = F}
select catergoryvariable, count(*)
from table 
group by 

```

case sensitive or case insensitive 

```{r, eval = F}

where variable ilike/like '%xxx%'

where lower(variable) = "xxx" 

where trim(" xx ") = "xx"
where rtrim(" xx ") = " xx"

```


split and concat text

```{r, eval = F}

select split_part(" a,b,c", ",", 2) = b


select concat("a", 2, "cc") = a2cc
```


create other class by using the `case when`

## Date and timestamps data

date/time data is not independent, it can be used in different way by modified it into a derived variable. 
we need to see it by units( seconds, minutes, hour, day, week, month, year), (UTC/global time stamps or the local time stamps) 
turn into a time series or find the gap would be help in further analysis.  

two time stamp format 

1. iso: international organization for standards: `YYYY-MM-DD HH:MM:SS`
2. UTC: Coordinated Universal Time `YYYY-MM-DD HH:MM:SS+HH`

```{r, eval = F}

Select date_part("month", date)

select date_trunc("month", date)


```
